---
title: "Data 607 - Assginment 1"
author: "Amit Kapoor"
date: "1/30/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Overview
This data originates from "Where People Go To Check The Weather". The source of the data is a Survey Monkey Audience poll commissioned by FiveThirtyEight and conducted from April 6 to April 10, 2015.


## Data Source
  https://fivethirtyeight.com/features/weather-forecast-news-app-habits/


## Data Description
* RespondentID
* Do you typically check a daily weather report? - Yes or No
* How do you typically check the weather? - The Weather Channel, Local TV News, Radio weather, Internet search, The default weather app on your phone, Newsletter, Newspaper, A specific website or app (please provide the answer)
* A specific website or app (please provide the answer)	If they responded this value for the second question, they were asked to write-in the app or website they used.
* If you had a smartwatch (like the soon to be released Apple Watch), how likely or unlikely would you be to check the weather on that device? - Very Likely, Somewhat Likely, Somewhat unlikely, Very unlikely
* Age - 18-29, 30-44, 45-59, 60+
* What is your gender? - Female, Male
* How much total combined money did all members of your HOUSEHOLD earn last year? - $0 to $9,999, $10,000 to $24,999, $25,000 to $49,999, $50,000 to $74,999, $75,000 to $99,999, $100,000 to $124,000, $125,000 to $149,999, $150,000 to $174,999, $175,000 to $199,999, $200,000+, Prefer not to answer.
* US Region - New England, Middle Atlantic, East North Central, West North Central, South Atlantic, East South Central, West South Central, Mountain, Pacific.



```{r data-weathercheck, echo=TRUE}
theLink <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/weather-check/weather-check.csv"

# load data into data frame
df_weathercheck <- read.csv(file=theLink, header = TRUE, sep = ",")

# display column names
colnames(df_weathercheck)
```

## Preview Data

```{r head-weathercheck, echo=TRUE}

# display header rows
head(df_weathercheck)
```


```{r total-rows, echo=TRUE}

# total num of rows
nrow(df_weathercheck)
```

```{r total-cols, echo=TRUE}

# total num of columns
ncol(df_weathercheck)
```

### Renaming Column names

```{r rename-colnames, echo=TRUE}

colnames(df_weathercheck)[2] <- "weather_chk"
colnames(df_weathercheck)[3] <- "weather_chk_src"
colnames(df_weathercheck)[4] <- "specific_src"
colnames(df_weathercheck)[5] <- "weather_chk_freq"
colnames(df_weathercheck)[7] <- "gender"
colnames(df_weathercheck)[8] <- "household_income"
colnames(df_weathercheck)[9] <- "us_region"

head(df_weathercheck)
```

### Summarizing data


```{r summary, echo=TRUE}

# display summary
summary(df_weathercheck[,-1])
```

## Subsetting

```{r subset, echo=TRUE}

df_weathercheck_subset <- subset(df_weathercheck, select = c(2,3,4,5,6,7))
head(df_weathercheck_subset)
```

### Plot between age and weather check frequency.

```{r plot1, echo=TRUE}

barplot(table(df_weathercheck_subset$Age, df_weathercheck_subset$weather_chk_freq), beside = TRUE, legend = TRUE)
```

### Filter records having columns weather_chk_freq and gender and preview them.

```{r filter-records, echo=TRUE}

# pipeline to filter
df_weathercheck_summary <- 
  subset(df_weathercheck, select = c('weather_chk_freq', 'gender')) %>% 
  group_by(gender) %>% 
  filter(!(weather_chk_freq == '-' & gender =='-')) 

head(df_weathercheck_summary)
```


### Plot between weather check frequency and gender.

```{r plot2, echo=TRUE}
barplot(table(df_weathercheck_summary$weather_chk_freq, df_weathercheck_summary$gender), beside = TRUE, legend = TRUE)
```



# Conclusions

I see there are 928 rows and 9 columns in this dataset. I renamed columns with shorter names for those who are too long. I did subset the data with 6 columns and generated couple of plots. Seeing the summary I conclude below.
* More females participated than males.
* Most of the participants typically check a daily weather report.
* Most of the participants use default weather app on their phone.

