---
title: "Data 607 - Assignment 3"
author: "Amit Kapoor"
date: "2/10/2020"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(dplyr)
library(stringr)
```



```{r data-college-majors, echo=TRUE}
theLink <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"

# load data into data frame
df_majors <- read.csv(file=theLink, header = TRUE, sep = ",")

head(df_majors)
```


```{r filter-majors, echo=TRUE}
df_majors %>% 
  filter(str_detect(Major, "DATA|STATISTICS"))
```

# Problem 2

```{r define-str, eval=TRUE}
given_str <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'

given_str
```


```{r extract-str, eval=TRUE}
# matches [a-z]+ for 1+ times
extracted_str <- str_extract_all(given_str, '[a-z]+\\s[a-z]+|[a-z]+')
extracted_str
```

```{r vec-str, eval=TRUE}
gsub('[\"]', '', extracted_str)
```


The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

#3 Describe, in words, what these expressions will match:

(.)\1\1

* (.) is first capturing group and . matches any character
* \1 matches the same text which is matched by 1st capturing grp
* \1 matches the same text which is matched by 1st capturing grp

(.)(.)\2\1

* (.) is first capturing group and . matches any character
* (.) is second capturing group and . matches any character
* \2 matches the same text which is matched by 2nd capturing grp
* \1 matches the same text which is matched by 1st capturing grp

(..)\1

* (..) is first capturing group where . matches any character and the next . matches any character
* \1 matches the same text which is matched by 1st capturing grp

(.).\1.\1

* (.) is first capturing group and . matches any character
* . matches any character
* \1 matches the same text which is matched by 1st capturing grp
* . matches any character
* \1 matches the same text which is matched by 1st capturing grp

(.)(.)(.).*\3\2\1

* " matches the character "
* (.) is first capturing group and . matches any character
* next (.) is second capturing group and . matches any character
* next (.) is third capturing group and . matches any character
* .* matches any character as many number of times
* \1 matches the same text which is matched by 3rd capturing grp
* \2 matches the same text which is matched by 2nd capturing grp
* \1 matches the same text which is matched by 1st capturing grp




#4 Construct regular expressions to match words that:

Start and end with the same character.

* ^(.).*\1$

```{r prb-4-1, eval=TRUE}
str.4.1 <- c("ZOO","YAML","SAS","TOY","MOM")  
pattern.4.1 <- "^(.).*\\1$"
str.4.1 %>% 
  str_subset(pattern.4.1)
```


Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

* (.)(.).*\1.

```{r prb-4-2, eval=TRUE}
str.4.2 <- c("church","papa","apple")  
pattern.4.2 <- "(.)(.).*\\1"
str.4.2 %>% 
  str_subset(pattern.4.2)
```


Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

* (.).\1.\1

```{r prb-4-3, eval=TRUE}
str.4.3 <- c("eleven","kjkjkjkj","apple")  
pattern.4.3 <- "(.).\\1.\\1"
str.4.3 %>% 
  str_subset(pattern.4.3)
```









