---
title: "Data607 - Project2C"
author: "Amit Kapoor"
date: "3/4/2020"
output:
  pdf_document: default
  html_document:
    toc: true
    toc_depth: 3
    theme: united
---


```{r include=FALSE}
library(dplyr)
library(tidyr)
library(kableExtra)
library(ggplot2)
```


# About the data

The data considered here is a dataset which has five colors cloths and their reaction to cooling and heating in every 10 minutes. Below are the variables description in the data

  * color - Cloth color
  * minute.0 - initial reaction 
  * minute.10 - temp after 10th minute (in Farenheit)
  * minute.20 - temp after 20th minute (in Farenheit)
  * minute.30 - temp after 30th minute (in Farenheit)
  * minute.40 - temp after 40th minute (in Farenheit)
  * minute.50 - temp after 50th minute (in Farenheit)
  * minute.60 - temp after 60th minute (in Farenheit)
  * phase - heating / colling
  
  
  
# Load the data


The dataset is placed in a csv file in the github repository and loaded here using read.csv function. kable library is used to display data tables.

```{r read-csv, echo=TRUE}

# get the github URL from github
theURL <- "https://raw.githubusercontent.com/amit-kapoor/data607/master/week6/hcdata.csv"

# read the data from csv
tempreact_df <- read.csv(theURL)

# show intial rows through head
kable(head(tempreact_df), align = "l")

```



# Tidying the data

Approach: My goal is to make data appear in a way which would be easy for further analysis. Here are the steps being performed.

* Rename columns where column names are abbreviated. Here I renamed all the columns starts with minute to make it meningful.
* Next is to filter out the records from original dataframe to separate records for cooling and heating. Here I created 2 dataframes respectively. I have used dplyr filter method to perform this operation.
* Also used gather function from tidyr to make data from wide to long for both cooling and heating datasets


```{r rename-cols, echo=TRUE}

# rename columns
tempreact_df <- tempreact_df %>% 
  rename("temp_initial"=minute.0) %>% 
  rename("temp_after_10"=minute.10) %>% 
  rename("temp_after_20"=minute.20) %>% 
  rename("temp_after_30"=minute.30) %>% 
  rename("temp_after_40"=minute.40) %>% 
  rename("temp_after_50"=minute.50) %>% 
  rename("temp_after_60"=minute.60)


# show intial rows through head
kable(head(tempreact_df), align = "l")

```

```{r}
cooling_phasedf <- filter(tempreact_df, tempreact_df$phase == "cooling")
kable(cooling_phasedf)
```


```{r cooling-long, echo=TRUE}
#Change from Wide to long
cooling_long_df <- gather(cooling_phasedf, Temperature, Value, "temp_initial":"temp_after_60")
cooling_long_df
```



```{r}
heating_phasedf <- filter(tempreact_df, tempreact_df$phase == "heating")
kable(heating_phasedf)
```


```{r heating-long, echo=TRUE}
#Change from Wide to long
heating_long_df <- gather(heating_phasedf, Temperature, Value, "temp_initial":"temp_after_60")
heating_long_df
```



# Data Analysis

For data analysis I used ggplot, geom_point, geom_line methods from ggplot2 library. I plotted the color on x-axis and temperature on y-axis. I created 2 graphs below to show the cooling and heating effects for different color in data.

```{r cooling, echo=TRUE}


cooling_phasedf %>% ggplot(aes(x=cooling_phasedf$color)) + scale_x_discrete(limits=cooling_phasedf$color) + 
  geom_point(aes(y=cooling_phasedf$temp_initial , group=1)) + geom_line(aes(y=cooling_phasedf$temp_initial, group=1, col="Intial Temp"), size=1) + 
  geom_point(aes(y=cooling_phasedf$temp_after_10, group=1)) + geom_line(aes(y=cooling_phasedf$temp_after_10, group=1, col="Temp after 10"), size=1) + 
    geom_point(aes(y=cooling_phasedf$temp_after_20 , group=1)) + geom_line(aes(y=cooling_phasedf$temp_after_20, group=1, col="Temp after 20"), size=1) + 
  geom_point(aes(y=cooling_phasedf$temp_after_30, group=1)) + geom_line(aes(y=cooling_phasedf$temp_after_30, group=1, col="Temp after 30"), size=1) + 
  geom_point(aes(y=cooling_phasedf$temp_after_40, group=1)) + geom_line(aes(y=cooling_phasedf$temp_after_40, group=1, col="Temp after 40"), size=1) +
  geom_point(aes(y=cooling_phasedf$temp_after_50, group=1)) + geom_line(aes(y=cooling_phasedf$temp_after_50, group=1, col="Temp after 50"), size=1) + 
  geom_point(aes(y=cooling_phasedf$temp_after_60, group=1)) + geom_line(aes(y=cooling_phasedf$temp_after_60, group=1, col="Temp after 60"), size=1) +
  labs(title="Cooling Temperature Versus Color", x="Color", y="Degrees in Farenheit", colour="") + 
  scale_colour_manual(values = c("orange", "grey" , "blue" , "green", "cyan" , "black" , "red"))
```



```{r heating, echo=TRUE}

heating_phasedf %>% ggplot(aes(x=heating_phasedf$color)) + scale_x_discrete(limits=heating_phasedf$color) + 
  geom_point(aes(y=heating_phasedf$temp_initial , group=1)) + geom_line(aes(y=heating_phasedf$temp_initial, group=1, col="Intial Temp"), size=1) + 
  geom_point(aes(y=heating_phasedf$temp_after_10, group=1)) + geom_line(aes(y=heating_phasedf$temp_after_10, group=1, col="Temp after 10"), size=1) + 
    geom_point(aes(y=heating_phasedf$temp_after_20 , group=1)) + geom_line(aes(y=heating_phasedf$temp_after_20, group=1, col="Temp after 20"), size=1) + 
  geom_point(aes(y=heating_phasedf$temp_after_30, group=1)) + geom_line(aes(y=heating_phasedf$temp_after_30, group=1, col="Temp after 30"), size=1) + 
  geom_point(aes(y=heating_phasedf$temp_after_40, group=1)) + geom_line(aes(y=heating_phasedf$temp_after_40, group=1, col="Temp after 40"), size=1) +
  geom_point(aes(y=heating_phasedf$temp_after_50, group=1)) + geom_line(aes(y=heating_phasedf$temp_after_50, group=1, col="Temp after 50"), size=1) + 
  geom_point(aes(y=heating_phasedf$temp_after_60, group=1)) + geom_line(aes(y=heating_phasedf$temp_after_60, group=1, col="Temp after 60"), size=1) +
  labs(title="Heating Temperature Versus Color", x="Color", y="Degrees in Farenheit", colour="") + 
  scale_colour_manual(values = c("orange", "grey" , "blue" , "green", "cyan" , "black" , "red"))
```

# Summary/Conclusion

This data used a small .csv file to demonstrate the color effect. I found above graphs interesting for both heating and cooling analysis. We can see above that black color cloth absorbs all colors and appears warmer than white color. More light a color absorbs, more warmer it would be.





